Manuscript 1 code

### 2024-09-13 manuscript_1_code.Rmd
figure 4 evenness site level done. code is messy as fuck we need to clean the markdown.

diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover))

@@

dss_ranges_site<- diversity_stability_synchrony_site %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability),
            min_synchrony = min(synchrony_trans),
            max_synchrony = max(synchrony_trans))
```

@@

>>>>

diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            evenness_mean = mean(evenness),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover))

@@

dss_ranges_site <- diversity_stability_synchrony_site %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_evenness = min(evenness_mean, na.rm = T),
            max_evenness = max(evenness_mean, na.rm = T),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability),
            min_synchrony = min(synchrony_trans),
            max_synchrony = max(synchrony_trans))
```

@@

# evenness stability site

## mod

```{r evenness synchrony mod}
even_stab_mod_site <- glmmTMB(cover_stability~ evenness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
summary(even_stab_mod_site)
Anova(even_stab_mod_site)

#                         Chisq Df Pr(>Chisq)   
# evenness_mean          0.1035  1   0.747623   
# habitat               11.8743  3   0.007827 **
# evenness_mean:habitat  9.6254  3   0.022034 *   
hist(residuals(even_stab_mod_site)) # blocky with some pos skew
plot(residuals(even_stab_mod_site) ~ fitted(even_stab_mod_site)) # heteroskedastic 
r.squaredGLMM(even_stab_mod_site) # 0.505
emtrends(even_stab_mod_site, pairwise ~ habitat, var = "evenness_mean") # no differences across habitats
```

even_stab_emm_site <- emmip(even_stab_mod_site,
                        habitat ~ evenness_mean,
                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
#make levels for habitat for plotting

filtered_even_stab_effects_site <- even_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)

#make levels for habitat for plotting
diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

```{r figure 4d}
(figure_4_evenness <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_even_stab_effects_site, aes(x = evenness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_even_stab_effects_site, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal evenness", title = "4f. Evenness site") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

ggsave(filename = "output/figure_4_evenness_site.png", figure_4_evenness, height = 10, width = 14)
```
